Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-31 introduction.tex
AA: piccole correzioni

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
The growing demand and reduced time-to-market of these systems are
introducing even more challenges that designers have to
face~\cite{wsts}.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, resulting in an intractable number
of simulations.

In these cases a Design Space Exploration (DSE) strategy is required,
i.e. an algorithm that selects only a tractable subset of all possible
configurations.  The final result of a design space exploration is a
subset of the configurations called \emph{Pareto set} (\cite{pareto},
see also \defR{pareto-set}). The corresponding values of the
objectives for each configuration of the Pareto Set is called
\emph{Pareto front}: it represents the trade-offs between objectives,
i.e.  for each point of the Pareto front there is no other
configuration performing better for all the objectives considered (see
\figR{psos}). Of course this is only a first high-level system
simulation, thus these few promising configurations will need to be
further evaluated in a next step of the design flow using a more
accurate low level simulation (e.g. register transfer level).

%while providing, at the end, configurations that are close to the
%optimum and that are worth evaluating, in a following step of the
%design process, in a more thorough way.

In this work we present PS, a multi-objective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front, i.e. how different these configurations are, in terms of objectives, with respect to the ones already found previously. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
The growing demand and reduced time-to-market of these systems are
introducing even more challenges that designers have to
face~\cite{wsts}.
Real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, resulting in an intractable number
of simulations.

In these cases a Design Space Exploration (DSE) strategy is required,
i.e. an algorithm that selects only a tractable subset of all possible
configurations.  The final result of a design space exploration is a
subset of configurations called \emph{Pareto set} (\cite{pareto},
see also \defR{pareto-set}). The set of the corresponding tuples of objective values is called \emph{Pareto front}: it represents the trade-off between objectives,
i.e.  for each point of the Pareto front there is no other
configuration performing better for all the objectives considered (see
\figR{psos}).
%while providing, at the end, configurations that are close to the
%optimum and that are worth evaluating, in a following step of the
%design process, in a more thorough way.

In this work we present PS, a multi-objective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front, i.e. how different these configurations are, in terms of objective values, with respect to the ones previously found. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there, rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

### 2014-10-31 paper.tex
AA: piccole correzioni

Embedded Systems Design requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters.
A simulation campaign can guide the design in finding the best
trade-offs, but due to the big number of possible configurations it
is often unfeasible to simulate all of them. For these reasons Design Space
Exploration (DSE) algorithms aim at finding near-optimal system
configurations only simulating a subset of them.

In this work we present PS, a new multi-objective optimization algorithm and evaluate it in the context of the exploration the embedded system design space. The basic idea is to recognize \emph{interesting} regions, i.e.
regions of the configuration space that are proving to provide better
configurations with regard to other ones. PS evaluates more configurations in
interesting regions while exploring less thoroughly the rest of the
configuration space. After a detailed formal description of the
algorithm and the underlying concepts, we show a 
case study involving the hardware/software exploration of a VLIW
architecture. Qualitative and quantitative comparisons of PS
against a well-known multi-objective genetic approach demonstrate that,
while not outperforming it in terms of pareto dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

>>>>

The Embedded Systems Design requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters.
A simulation campaign can guide the design in finding the best
trade-offs, but, due to the big number of possible configurations, it
is often unfeasible to simulate them all. For these reasons, Design Space
Exploration (DSE) algorithms aim at finding near-optimal system
configurations, by simulating only a subset of them.

In this work we present PS, a new multi-objective optimization algorithm and evaluate it in the context of the embedded system design. The basic idea is to recognize \emph{interesting} regions, i.e.
regions of the configuration space that are proving to provide better
configurations with respect to other ones. PS evaluates more configurations in the
interesting regions, while less thoroughly exploring the rest of the
configuration space. After a detailed formal description of the
algorithm and the underlying concepts, we show a 
case study involving the hardware/software exploration of a VLIW
architecture. Qualitative and quantitative comparisons of PS
against a well-known multi-objective genetic approach demonstrate that,
while not outperforming it in terms of Pareto dominance, the proposed
approach can balance the uniformity and granularity qualities of the
solutions found, obtaining more extended Pareto-fronts that provide a wider view of the potentiality of the designed device. Therefore, PS represents a
further valid choice for the designer, when objective constrains allow it.
\end{abstract}
